Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test

### 2024-04-23 manuscript/manuscript.tex
many edits

The standard one-sample Kolmogorov-Smirnov (KS) test is widely
recognized as an effective good-of-fit test for continuous distributions.
Consider $X_1, ..., X_n$, a random sample of size~$n$ from some continuous
distribution and the null hypothesis $H_0$ that $X_i$'s follow a specific
hypothesized distribution~$F$.
If we let $F_n(t) = \sum_{i=1}^n I(X_i \le t) / n$ be the empirical cumulative
distribution function of the sample, where $I(\cdot)$ is the indicator
function, the KS test statistic takes the form
\begin{equation}
  \label{eq:ks_standard}
  T_n = \sqrt{n} \sup_x | F_{n}(x) - F(x) |.
\end{equation}
As sample size $n\to \infty$, the distribution of $T_n$ converges to that of the
absolute value of standard Brownian bridge, which is known as the Kolmogorov
distribution \citep{stephens1974edf}. This distribution function is
precisely computable using contemporary statistical software
\citep{marsaglia2003evaluating}. The versatility of the KS test allows its
application across various domains, such as analyzing cosmic microwave
background radiation \citep{naess2012application}, monitoring the count rate of
radioactive data \citep{aslam2020introducing}, gear condition monitoring
\citep{andrade2001gear}, studying images of breast cancer tumors
\citep{demidenko2004kolmogorov}, and examining financial markets
\citep{lux2001turbulence}.

@@

Another prevalent misapplication of the KS test arises when data exhibit serial
dependence, which is often overlooked in analyses. For example, in an
investigation of performance variation in high-performance computing systems,
\citet{tuncer2019ieee} did not describe how or if they accounted for serial
dependence when applying the two-sample KS test.
\citet{zeimbekakis2022misuses} proposed a semi-parametric bootstrap
where the serial dependence structure is modeled via a working serial copula.
A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent. For this,
\citet{zeimbekakis2022misuses} also recommends a semi-parametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for.

@@

\jy{State clearly the test statistic definition. Point out that the null
  distribution depends on the unknown parameters and unknown serial
  dependence. Point out the challenges are to approximate the null distribution.}

We note that
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x)) - 
\sqrt{n}(F(x; \hat\theta_n) - F(x)).
\end{equation*}
\jy{clean up the notations. Has $F$ been defined yet? do you need cdf?}
\mc{addressed}

@@

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$ with Kendall's
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it does not vary between two different 
marginal distributions.
To generate the samples to which our
method would be applied, we simulated a time series $W_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
W_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: 
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$, as these
correspond to the desired values for $\tau$. First, we generated a
marginal $N(8, 8)$ by marginally transforming $W_t$ by
\begin{equation*}
X_t = F^{-1}[\Phi(W_t)],
\end{equation*}
\jy{define $\Phi$}\mc{addressed}
\jy{The language issues are more than I expected. Please do a careful proofread,
  now that you are more experienced.}
where $F^{-1}(p)$ is the quantile function for the $N(8, 8)$ 
distribution and $\Phi$ is the distribution function of the standard normal
distribution.
Then we generated a marginal gamma series by the same procedure, but
replacing $F^{-1}(p)$ with the quantile function for the $\Gamma(8, 1)$
distribution.
After the transformation
to the marginal gamma distribution, the autocorrelations are $\{-0.884, 
-0.686, -0.389, -0.033, 0.327, 0.639, 0.861\}$.

These distributions
were chosen to compare results on normal and non-normal
error structures. The specific parameters were chosen because the distributions 
are very similar and their
first two moments are the same. 
For the purposes of 
evaluating if the test holds it size, when $X_t \sim N(8, 8)$, we tested that the 
marginal distribution is from the
Normal family, or 
$X_t \sim N(\cdot \mid \mu, \sigma^2), \quad t = 1, \ldots, n$
for some unspecified $\mu$ and $\sigma$. When $X_t \sim \Gamma(8, 1)$, we tested
that the marginal distribution family is from the Gamma family,
or
$X_t \sim \Gamma(\cdot \mid \alpha, \beta), \quad t = 1, \ldots, n$.
for some unspecified $\alpha$ and $\beta$.
For the block
bootstrap step,
we used $B = 1000$ and $l = \lceil n^{1/3} \rceil$.
For each setting for $F$, $\tau$, and $n$, we replicate the method 1000 times 
to get the distribution of the p-values 
for the test when applied to samples from the same data generating process.

Using the \textsl{qqplotr} and \textsl{ggplot2} packages 
\citep{qqplotr, ggplot2},
we constructed Q-Q plots of the distribution of the p-values to see if they are
uniformly distributed. If the distribution of the p-values is uniform, this 
indicates that under the null hypothesis, our method will only reject the $H_0$
at a rate of $\alpha$.
A zoomed in plot for probabilities between 0 and 0.1 is
also provided, as
that is the most common range for significance levels $\alpha$.

@@

\input{tables/rr_100}

\input{tables/rr_200}

\input{tables/rr_400}

\input{tables/rr_800}

From Figures~\ref{fig:normal} and \ref{fig:gamma}, we can observe that 
although the extent to which the p-values appear to be 
completely aligned with the line
in the Q-Q plots is dependent on the sample size and the level of serial 
dependence.
A small sample size like $n = 100$ or $200$ seems adequate when Kendall's
$\tau$ is lower than 0.25. For Kendall's $\tau \geq 0.5$, a sample larger than
$n = 200$ seems necessary for the p-values to be normally distributed. For
Kendall's $\tau \geq 0.75$, a sample larger than $n = 800$ appears to be 
necessary, as the p-values are not aligned with the line in the Q-Q plots. This
is not necessarily a cause for concern, as (for normal margins)
a Kendall's $\tau$ corresponds to
a $\phi$ of 0.924, which is very high. Additionally, performance generally seems
better for negative $\tau$ values versus positive $\tau$ values of the same
strength. Results are very similar for normal and gamma margins.
\jy{Summarize findings in bullet points before forming text.}

Under $H_0$, the p-values for the non-parametric block bootstrap KS test are
\jy{Not true from the plots.}\mc{addressed}
uniformly distributed as long as the serial dependence is not too high. 
This is an indication that our method 
holds its size for moderate serial dependence levels.
For such cases, we can expect the probability
of a type 1 error to be the significance level $\alpha$.

\subsection{Power}
The power of the proposed test is investigated with data generated from
distributions that are not in the hypothesized distribution family. If the test
is indeed powerful, we would ideally want the probability of
rejecting $H_0$ when $H_0$ is false\jy{beta has been
  used. Discuss in terms of power: the probability of rejecting $H_0$ when $H_0$
  is false.}
to be 1, but we
expect it to be generally high. Again, we must try the method with different
marginal distributions to ensure that it is robust.
We would then expect the distribution of p-values to be non-uniform, and the 
rate \jy{Does not sound right. What is $\alpha$?}
of $\#P < \alpha$ to be high, meaning a higher density of low p-values.

\jy{Description too wordy, Be concise. Basically generate data from one
  distribution but test for the other distribution.}
Like we did to see if the test holds its size, we generated time series with 
marginal distributions $N(8, 8)$ and 
$\Gamma(8, 1)$, Kendall's $\tau \in \{-.75, -.5, -.25, 0, .25, .5, .75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. However, 
to evaluate the test's power, when $X_t \sim N(8, 8)$, we tested
that the marginal distribution family is from the Gamma family,
or
$X_t \sim \Gamma(\cdot \mid \alpha, \beta), \quad t = 1, \ldots, n$.
for some unspecified $\alpha$ and $\beta$. Because the support of $N(8, 8)$ is
$(-\infty, \infty)$, but the support of $\Gamma(8, 1)$ is $(0, \infty)$, we used
the \textsl{truncdist} package \citep{truncdist} to truncate the series at 
values less than or equal to 0.
When $X_t \sim \Gamma(8, 1)$, we tested that the 
marginal distribution is from the
Normal family, or 
$X_t \sim N(\cdot \mid \mu, \sigma^2), \quad t = 1, \ldots, n$
for some unspecified $\mu$ and $\sigma$. For the block bootstrap step,
we used $B = 1000$ and $l = \lceil n^{1/3} \rceil$.
For each setting for $F$, $\tau$, and $n$, we replicate the method $1000$ times 
to obtain a p-value $p_r$ for each $r \in \{1, \ldots, 1000\}$.

\jy{alpha has been used as gamma parameters. This is not necessary, Just say we
  checked the power when the significance level is .05.}
Given some significance level $\alpha$, we can compute a rejection rate 
$q = \#\{p_r < \alpha\} / 1000$ for $r \in \{1, \ldots, 1000\}$.
Additionally, we can construct a 95\% confidence interval for 
this rate. Below are tables containing rejection rates showcasing the 
differences when simulation settings are changed.

@@

\jy{Discuss the difference between the semiparametric bootstrap and the proposed
  nonparametric bootstrap. The semiparametric method depends on a working copula
  model for the serial dependence. If that is severely misspecified, I'd expect
  poor performance. The proposed method does not need to specify that dependence
  structure.}

@@

\jy{Stock returns are known to be serially dependent and heavy-tailed. Search
  google scholar ``stylized facts of daily returns series'' and cite them.}

@@

The same analyses were repeated on the daily returns of the S\&P 100 index.
Our method demonstrates more instances where it reaches
different conclusions at the 0.05 significance level compared to other methods.
Our method does not reject the hypothesis that the series conforms to a
Student's $t$ distribution with degrees of freedom $v = 5$, yielding a p-value
of 0.0873. In contrast, \citet{babu2004goodness}'s method yields a p-value of 
0.0368, and \citet{zeimbekakis2022misuses}'s method yields a p-value of 0.03.
conclusions regarding the marginal distribution of the population under scrutin.

@@

We have also 
demonstrated possible applications of this method to financial data.
\jy{don't astrophie citet; let alone citep}\mc{addressed}
Through comparison with the non-parametric
bootstrap bias correction developed by \citet{babu2004goodness}
and the parametric
bootstrap method developed by \citet{zeimbekakis2022misuses}, we have 
underscored the importance of accounting for serial
dependence in choosing an appropriate methodology.
This method holds relevance for studies aiming to assess if a serially dependent
time series adheres to a hypothesized distribution in scenarios where parameters
are unknown or unspecified. Future studies could investigate
\jy{Too vague.}
the method's performance on different marginal distributions and
different dependence structures. Additionally, one could propose a 
similar method
for comparing two serially dependent methods. One could also apply the method
showcased
in this paper to fields like earthquake prediction and astronomy.

>>>>

The standard one-sample Kolmogorov-Smirnov (KS) test is widely
recognized as an effective good-of-fit test for continuous distributions.
Consider $X_1, ..., X_n$, a random sample of size~$n$ from some continuous
distribution and the null hypothesis $H_0$ that $X_i$'s follow a specific
hypothesized distribution~$F(x; \theta_0)$, where $\theta_0$ are the 
specified parameters
of the hypothesized distribution family.
If we let $F_n(t) = \sum_{i=1}^n I(X_i \le t) / n$ be the empirical cumulative
distribution function of the sample, where $I(\cdot)$ is the indicator
function, the KS test statistic takes the form
\begin{equation}
  \label{eq:ks_standard}
  T_n = \sqrt{n} \sup_x | F_{n}(x) - F(x; \theta_0)|.
\end{equation}
As sample size $n\to \infty$, the distribution of $T_n$ converges to that of the
absolute value of standard Brownian bridge, which is known as the Kolmogorov
distribution \citep{stephens1974edf}. This distribution function is
precisely computable using contemporary statistical software
\citep{marsaglia2003evaluating}. The versatility of the KS test allows its
application across various domains, such as analyzing cosmic microwave
background radiation \citep{naess2012application}, monitoring the count rate of
radioactive data \citep{aslam2020introducing}, gear condition monitoring
\citep{andrade2001gear}, studying images of breast cancer tumors
\citep{demidenko2004kolmogorov}, and examining financial markets
\citep{lux2001turbulence}.

@@

Another prevalent misapplication of the KS test arises when data exhibit serial
dependence, which is often overlooked in analyses. For example, in an
investigation of performance variation in high-performance computing systems,
\citet{tuncer2019ieee} did not describe how or if they accounted for serial
dependence when applying the two-sample KS test.
The distribution of the test statistic under
the null hypothesis depends on the unknown parameters and unknown 
serial dependence. Therefore, the null distribution under dependence cannot be
used,
and the challenge is to approximate said distribution.
\citet{zeimbekakis2022misuses} proposed a semi-parametric bootstrap
where the serial dependence structure is modeled via a working serial copula.
A particularly complex scenario arises when the distribution has unspecified
parameters, and the data are serially dependent. For this,
\citet{zeimbekakis2022misuses} also recommends a semi-parametric bootstrap
approach, where the parameters of the hypothesized distribution and the serial
dependence both need to be estimated and their uncertainty accounted for,
but this could affect results if the estimated serial
dependence is too far from the truth. For this reason, a non-parametric method
is motivated.

@@

\jy{State clearly the test statistic definition. Point out that the null
  distribution depends on the unknown parameters and unknown serial
  dependence. Point out the challenges are to approximate the null distribution.}
\mc{addressed in third paragraph of introduction}

We note that
\begin{equation*}
Y_n(x; \hat\theta) = \sqrt{n}(F_n(x) - F(x)) - 
\sqrt{n}(F(x; \hat\theta_n) - F(x)),
\end{equation*}
\jy{clean up the notations. Has $F$ been defined yet? do you need cdf?}
\mc{addressed}
where $F(x)$ is the true cumulative distribution function 
(under the $H_0: F(x) = F(x; \theta_0)$ for some true
parameter $\theta_0$.)

@@

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$ with Kendall's
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it does not vary between two different 
marginal distributions.
To generate the samples to which our
method would be applied, we simulated a time series $W_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
W_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: 
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$, as these
correspond to the desired values for $\tau$. First, we generated a
marginal $N(8, 8)$ by marginally transforming $W_t$ by
\begin{equation*}
X_t = F^{-1}[\Phi(W_t)],
\end{equation*}
\jy{define $\Phi$}\mc{addressed}
\jy{The language issues are more than I expected. Please do a careful proofread,
  now that you are more experienced.}
where $F^{-1}(p)$ is the quantile function for the $N(8, 8)$ 
distribution and $\Phi$ is the distribution function of the standard normal
distribution.
Then we generated a marginal gamma series by the same procedure, but
replacing $F^{-1}(p)$ with the quantile function for the $\Gamma(8, 1)$
distribution.
After the transformation
to the marginal gamma distribution, the autocorrelations are $\{-0.884, 
-0.686, -0.389, -0.033, 0.327, 
\break 0.639, 0.861\}$.

These distributions
were chosen to compare results on normal and non-normal
error structures. The specific parameters were chosen because the distributions 
are very similar and their
first two moments are the same. 
For the purposes of 
evaluating if the test holds it size, when $X_t \sim N(8, 8)$, we tested that the 
marginal distribution is from the
Normal family, or 
$X_t \sim N(\cdot \mid \mu, \sigma^2), \quad t = 1, \ldots, n$
for some unspecified $\mu$ and $\sigma$. When $X_t \sim \Gamma(8, 1)$, we tested
that the marginal distribution family is from the Gamma family,
or
$X_t \sim \Gamma(\cdot \mid \alpha, \beta), \quad t = 1, \ldots, n$.
for some unspecified $\alpha$ and $\beta$.
For the block
bootstrap step,
we used $B = 1000$ and $l = \lceil n^{1/3} \rceil$.
For each setting for $F$, $\tau$, and $n$, we replicate the method 1000 times 
to get the distribution of the p-values $p_r$
for the test when applied to samples from the same data generating process.

Using the \textsl{qqplotr} and \textsl{ggplot2} packages 
\citep{qqplotr, ggplot2},
we constructed Q-Q plots of the distribution of the p-values to see if they are
uniformly distributed. If the distribution of the p-values is uniform, the
points will be aligned with the diagonal line. This 
indicates that under the null hypothesis, our method will only reject the $H_0$
at the rate of the significance level.

@@

\input{tables/rr_norm}

\input{tables/rr_gamma}

From Figures~\ref{fig:normal} and \ref{fig:gamma}, we can observe that 
the extent to which the p-values are uniformly distributed
is dependent on the sample size and the level of serial 
dependence.
A small sample size like $n = 100$ or $200$ seems adequate when Kendall's
$\tau$ is lower than 0.25. For Kendall's $\tau \geq 0.5$, a sample larger than
$n = 200$ seems necessary for the p-values to be normally distributed. For
Kendall's $\tau \geq 0.75$, a sample larger than $n = 800$ appears to be 
necessary, as the p-values are not aligned with the line in any of the Q-Q 
plots. This
is not necessarily a cause for concern, as (for normal margins)
a Kendall's $\tau$ corresponds to
a $\phi$ of 0.924, which is very high.
Additionally, positive serial dependence appears to misalign the distribution
of the p-values moreso than negative serial dependences of the same strength.
Results are very similar for normal and gamma margins.
\jy{Summarize findings in bullet points before forming text.}

Under $H_0$, the p-values for the non-parametric block bootstrap KS test are
\jy{Not true from the plots.}\mc{addressed}
uniformly distributed as long as the serial dependence is not too high. 
This is an indication that our method 
holds its size for moderate serial dependence levels.
For such cases, we can expect the probability
of a type 1 error to be approximately equivalent to the significance level.

From Tables~\ref{table:rr_norm} and \ref{table:rr_gamma}, we can observe 
that the rejection rate is close to the significance level
as long as 
Kendall's $\tau$ is less than or equal to 0.5 and a sample size greater
than 100 is used.

\subsection{Power}
The power of the proposed test is investigated with data generated from
distributions that are not in the hypothesized distribution family. If the test
is indeed powerful, we would ideally want the probability of
rejecting $H_0$ when $H_0$ is false\jy{beta has been
  used. Discuss in terms of power: the probability of rejecting $H_0$ when $H_0$
  is false.}\mc{addressed}
to be 1, but we
expect it to be generally high. Again, we must try the method with different
marginal distributions to ensure that it is robust.
We would then expect the distribution of p-values to be non-uniform -
specifically a higher density of low p-values - and the 
rate \jy{Does not sound right. What is $\alpha$?}\mc{addressed}
of $\#(p_r < 0.05)$ - where 0.05 is the significance level - to be high,

\jy{Description too wordy, Be concise. Basically generate data from one
  distribution but test for the other distribution.}\mc{addressed}
We generated the same scheme of time series as we did when testing size, but
tested that the time series was from the other distribution.
Because the support of $N(8, 8)$ is
$(-\infty, \infty)$, but the support of $\Gamma(8, 1)$ is $(0, \infty)$, we used
the \textsl{truncdist} package \citep{truncdist} to truncate the series at 
values less than or equal to 0 when testing that $N(8, 8)$ sample was
from $\Gamma(8, 1)$. For each setting for $F$, $\tau$, and $n$, we replicate the 
method $1000$ times 
to obtain a p-value $p_r$ for each $r \in \{1, \ldots, 1000\}$.

\jy{alpha has been used as gamma parameters. This is not necessary, Just say we
  checked the power when the significance level is .05.}\mc{addressed}
We can compute a rejection rate 
$q = \#\{p_r < 0.05\} / 1000$ for $r \in \{1, \ldots, 1000\}$.
Additionally, we can construct a 95\% confidence interval for 
this rate. Below are tables containing rejection rates showcasing the 
differences when simulation settings are changed.

@@

\jy{Discuss the difference between the semiparametric bootstrap and the proposed
  nonparametric bootstrap. The semiparametric method depends on a working copula
  model for the serial dependence. If that is severely misspecified, I'd expect
  poor performance. The proposed method does not need to specify that dependence
  structure.}
\mc{addressed earlier}

@@

\jy{Stock returns are known to be serially dependent and heavy-tailed. Search
  google scholar ``stylized facts of daily returns series'' and cite them.}
\mc{addressed}

@@

The same analyses were repeated on the daily returns of the S\&P 100 index.
Our method demonstrates more instances where it reaches
different conclusions at the 0.05 significance level compared to other methods.
Our method does not reject the hypothesis that the series conforms to a
Student's $t$ distribution with degrees of freedom $v = 5$, yielding a p-value
of 0.0873. In contrast, \citet{babu2004goodness}'s method yields a p-value of 
0.0368, and \citet{zeimbekakis2022misuses}'s method yields a p-value of 0.03.
conclusions regarding the marginal distribution of the population under 
scrutiny.

@@

We have also 
demonstrated possible applications of this method to financial data.
\jy{don't astrophie citet; let alone citep}\mc{addressed}
Through comparison with the non-parametric
bootstrap bias correction developed by \citet{babu2004goodness}
and the parametric
bootstrap method developed by \citet{zeimbekakis2022misuses}, we have 
underscored the importance of accounting for serial
dependence in choosing an appropriate methodology.
This method holds relevance for studies aiming to assess if a serially dependent
time series adheres to a hypothesized distribution in scenarios where parameters
are unknown or unspecified. Future studies could investigate
\jy{Too vague.}\mc{addressed}
the method's performance on different marginal distributions such
as the generalized extreme value distribution and
different dependence structures such as MA, ARMA, and ARIMA processes. Additionally, one could propose a 
similar method
for comparing two serially dependent methods. One could also apply the method
showcased
in this paper to fields like earthquake prediction and astronomy.
